---
title: "Interacting with time"
author: "Keegan Gunderson"
date: "March 11, 2018"
output: 
  html_document: 
    keep_md: yes
    code_folding: hide
---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(tidyquant)
library(quantmod)
library(lubridate)
library(DT)
library(timetk)
library(dygraphs)
```


```{r}
#get stock ticker
tickers <- c("CXW", "F", "GM", "JCP", "KR", "WDC", "NKE","T", "WDAY", "WFC", "WMT")

#for ggplot
stock1 <- tq_get(tickers, get  = "stock.prices",
        from = "2013-1-01",
         to   = "2018-1-01")

#get that data for the last 5 years for dygraph
stock2 <- tq_get(tickers, get  = "stock.prices",
         from = "2013-1-01",
         to   = "2018-1-01") %>%
  select(symbol, date, adjusted) %>%
  spread(key = symbol, value = adjusted) %>%
  tk_xts(silent = TRUE)
  
```

```{r}
#Interactive graph
dygraph(stock2, main = "Stock Price for last 5 Years", 
        ylab = "Share Price", 
        xlab = "Daily Prices") %>% 
  dyHighlight(highlightCircleSize = 3,
              highlightSeriesBackgroundAlpha = .2,
              hideOnMouseOut = FALSE,
              highlightSeriesOpts = list(strokeWidth = 3)) %>% 
  dyLegend(show = "follow") %>% 
  dyRangeSelector()

```

Summarizing my interactive graphic:  I added a range selector so you could zoom in and see specific share prices each day if you wanted.  I added the legend to follow you along as you move through the graph, I realize that this can be annoying so I am thinking about changing it. I also highlighted the line width so when you hover over it you are able to see the color and line better. I think that this turned out decent.


```{r}
#ggplot cant choose one but I think the col looks better
stock3 <- stock1 %>%
  mutate(month = floor_date(date, "month")) %>% 
  group_by(symbol, month) %>%
  summarise(volume = sum(volume))

ggplot(data = stock3, aes(x = month, y = volume/10000000, color = symbol)) +
  geom_col(colour = "black",
           fill = "#8000ff") +
  labs(title = "Company Stock Volumes",
       subtitle = "each bar is a month",
       x = "Date",
       y = "Stock Volume in millions",
       color = "Company") +
  facet_grid(symbol~.)

ggplot(data = stock3, aes(x = month, y = volume/10000000, color = symbol)) +
  geom_point() +
  labs(title = "Company Stock Volumes",
       subtitle = "each dot is a month",
       x = "Date",
       y = "Stock Volume in millions",
       color = "Company") +
  facet_grid(symbol~.)

```

Summarizing my ggplot graphic:  I have two different graphs that show the exact same information because I was not sure which one I liked better, but I am leaning towards to geom_col graph because it is pretty easy to read. I am thinking about making the x axis ticks in months instead of just years to clarify that each column or dot is a month. You should be able to tell that this graph is for company stock volumes, overall I think this turned out okay. I need to mess with it and make it look a little better, comment in mine with suggestions to do please. Should I keep my bars all one color or make them different per group?